---
title: Introduction to dplyr
author: Gaurav Satav
date: '2018-03-01'
slug: introduction-to-dplyr
categories: []
tags:
  - notes
  - learning
---

This is a short introduction to the r package `dplyr`. The datasets we'll be using for the purposes of this tutorial will are the ones present within the `datasets` library of R.

Lets first start by importing the required libraries.
```{r,echo=FALSE}
library(knitr)
library(kableExtra)
```


```{r Importing Libraries,message=FALSE}
library(datasets)
library(dplyr)
```

The best way to appreciate `dplyr` is to first get a feel of why is it required in the first place. In order to do that, lets say, given the data below, 

```{r,echo=FALSE}
head(mtcars)
```
we'd like to perform following actions on the dataframe.

1. Create a new dataframe which consist of only carnames,mpg and cyl

2. Create a new dataframe which consist only of cyl value = 4

and would like to obtain answers to the following questions.

1. 


The most common set of actions which we'll be performing would be

* Select particular Variables (Columns)

* Selecting particular Observations (Rows)

* Grouping Observations.

* Summarize variables

* Creating new dataframes.



## Grouping


## Summarizing 

*dplr* offers us with the following list of summarising functions:

* summarise()

* summarise_all()

* summarise_each()

* summarise_at()

* summarise_if()

Hmm, now what's the purpose of each of these similar looking functions? Well we can start of by getting an idea of what the output of a summarise function is.


Now suppose we would like to know, what is the mean (this is our function) **mpg** value of all the 4,6 and 8 cylinder (cyl is our variable) cars. Using the summarise function the code to generate output would look as follows

```{r}
mtcars %>% group_by(Cylinders = cyl) %>% summarise("MEAN MPG"= mean(mpg))
```

as can be seen from the output above, the mean **mpg** value for a 4 cyl vehicle is 26.7

Suppose along with finding the mean, we'd like to find the min value of for each cyl value. In this case we have to summarise our data based on 2 functions, **mean** and **min**. Our variable remains the same (mpg). The code would look as follows

```{r,}
mtcars %>% group_by(Cylinders = cyl) %>% summarise_at(vars(mpg,disp),.funs = c(mean,min))
```

**How many variables do I need to summarise?** 

**How many different functions do I need the columns to be summarised by?**


